---
title: "SampleFields R-package"
author: "Fabian Telschow"
date: "01/28/2021"
output: pdf_document
---

  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=6, fig.height=3)
library( SampleFields )
library( plotly )
library( tidyverse )
```

# 1. Some guidelines through the package
## 1.1. Package dependencies
At the moment the package depends on the following R-packages:

* aws (developed by Karsten Tabelow and Jörg Pohlzehl): Used for fast smoothing
to obtain random fields up to dimension 3.
* tidyverse: mainly to use ggplot
* plotly: optional for "interactive" ggplots

## 1.2. Design of the package
The main functions of the package are contained in the file ***DataModels.R***
currently only the signal-plus-noise model is contained, i.e.,
$$
  Y(x) = \mu(x) + \sigma(x) Z(x)\,.
$$
Here $\mu(x)$ is the population mean function, $\sigma(x)$ the population
standard deviation function and $Z$ a random error process. Predefined functions
for each of these ingredients can be found in the folders ***MeanFunctions.R***,
***SDFunctions.R*** and ***ErrorFields.R***.

Population mean functions and
standard deviations can be easily defined as R-functions which take a vector of
locations as input and output the function values on the input coordinates x.
Note that locations are assumed to be matrices where ***ROWS*** define the
coordinates. Default values are that the population mean is zero and the
standard deviation 1 for all $x$. A simple 1D examples is
```{r}
# Sample size
N = 5

# 1D locations
x = seq( 0, 2, length.out = 100 )

# define mean function
mu <- function( x ){
  3*x
} 

# Get 10 samples with mean mu, std=1 and Z a default sum of random Gaussians
Y <- SignalPlusNoise( N = N, x = x, mu = mu )

# plot the sample
plot( Y )
```

Similarly in 2D we can produce samples by:

```{r}
# Sample size
N = 6

# 1D locations
x = seq( 0, 2, length.out = 70 )
# make the vector a grid of 2D coordinates
x = expand.grid( x, x )

# define mean function
mu <- function( x ){
  3*x[,1]*x[,2]
} 

# Get 10 samples with mean mu, std=1 and Z a default sum of random Gaussians
Y <- SignalPlusNoise( N = N, x = x, mu = mu )

# plot the sample
plot( Y, ncols = 2 )
```


## 1.2. RandomField class
The output of *SignalPlusNoise()* is a class object. I called this class "RandomField". An S3 class is basically a list object. The list contains the follwing entries
```{r}
names( Y )
```

* **values** is a nloc x N matrix containing in each column a realization of the random field.
* **locations** is the vector/matrix containing the coordinate locations of the random field
* **dim** contains [nloc, N]
* **D** is the dimension of the domain
* **nloc** is the number of locations, i.e., the number of rows in $x$
* **N** is the number of realizations of the random field.

To implement your own Noise function please look into ***ErrorFields.R*** here you can find examples and how to construct the S3 calss object.

## 1.3. Observation noise
The signal-plus-noise model allows also to add observation noise.  The corresponding model is
$$
  Y(x_i) = \mu(x_i) + \sigma(x_i) Z(x_i) + \epsilon_i\,.
$$
Here $x_i$ is an observed location and $\epsilon_i$ is a sample of some distribution. By default this noise is i.i.d. Gaussian with a standard deviation which can be specified.

You can generate samples from this model as follows:
```{r}
# Sample size
N = 5

# 1D locations
x = seq( 0, 2, length.out = 100 )

# define mean function
mu <- function( x ){
  3*x
} 

# Get 10 samples with mean mu, std=1 and Z a default sum of random Gaussians abd
# iid Gaussian observation noise with sd = 0.5.
Y <- SignalPlusNoise( N = N, x = x, mu = mu, obs.noise = 0.5 )

# plot the sample
plot( Y, main = "Sample with added observation noise" )
```

If you want to define say depend error processes or observation noise from different distributions you can do it as follows
```{r}
# Define a modified sampling function
rate = 2;
rexp.mod <- function( N, x ){
  tmp = getDim( x )
  return( matrix( rexp( tmp$nloc * N, rate ), tmp$nloc, N ) )
}

# Get 10 samples with mean mu, std=1 and Z a default sum of random Gaussians abd
# iid Gaussian observation noise with sd = 0.5.
Y <- SignalPlusNoise( N = N, x = x, mu = mu, obs.noise = rexp.mod )

# plot the sample
plot( Y, main = "Sample with added observation noise" )
```


## 1.3. Things to do
* maybe implement different mean, standard deviation or noise processes
* some functions need more documentation and an example section would be great.
* we should discuss what we need for the plotting function and implement it.
* maybe you have suggestions for better usability.
